High-Dimensional Regularized Discriminant Analysis

### 2013-05-07 reports/paper-rda.Rnw
Last minute tweaks. Submitted
Removed Biometrika LaTeX template from paper. However, I have not yet added the bolding back.

\documentclass[lineno]{biometrika}

\usepackage{graphicx, amsmath, amssymb, bm, natbib}

\newtheorem{cor}{Corollary}
\newtheorem{thm}{Theorem}

@@

%% Please use the following statements for
%% managing the text and math fonts for your papers:
\usepackage{times}
%\usepackage[cmbold]{mathtime}

\usepackage[plain,noend]{algorithm2e}

\makeatletter
\renewcommand{\algocf@captiontext}[2]{#1\algocf@typo. \AlCapFnt{}#2} % text of caption
\renewcommand{\AlTitleFnt}[1]{#1\unskip}% default definition
\def\@algocf@capt@plain{top}
\renewcommand{\algocf@makecaption}[2]{%
  \addtolength{\hsize}{\algomargin}%
  \sbox\@tempboxa{\algocf@captiontext{#1}{#2}}%
  \ifdim\wd\@tempboxa >\hsize%     % if caption is longer than a line
    \hskip .5\algomargin%
    \parbox[t]{\hsize}{\algocf@captiontext{#1}{#2}}% then caption is not centered
  \else%
    \global\@minipagefalse%
    \hbox to\hsize{\box\@tempboxa}% else caption is centered
  \fi%
  \addtolength{\hsize}{-\algomargin}%
}
\makeatother

%%% User-defined macros should be placed here, but keep them to a minimum.
\def\Bka{{\it Biometrika}}
\def\AIC{\textsc{aic}}
\def\T{{ \mathrm{\scriptscriptstyle T} }}
\def\v{{\varepsilon}}

\newcommand{\xbar}{\bar{x}}
\newcommand{\tr}{\text{tr}}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

\jname{Biometrika}
%% The year, volume, and number are determined on publication
\jyear{2013}
\jvol{99}
\jnum{1}
%% The \doi{...} and \accessdate commands are used by the production team
%\doi{10.1093/biomet/asm023}
\accessdate{Advance Access publication on 31 July 2012}
\copyrightinfo{\Copyright\ 2013 Biometrika Trust\goodbreak {\em Printed in Great Britain}}

%% These dates are usually set by the production team
\received{April 2013}
\revised{April 2013}

%% The left and right page headers are defined here:
\markboth{J. A. Ramey, C. K. Stein \and D. M. Young}{High-Dimensional
  Regularized Discriminant Analysis}

%% Here are the title, author names and addresses
\title{High-Dimensional Regularized Discriminant Analysis}

% \author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
\author{J. A. RAMEY}
\affil{Vaccine and Infectious Disease Division, Fred Hutchinson Cancer Research
  Center,\\ Seattle, Washington 98115, U.S.A. \email{johnramey@gmail.com}}

\author{C. K. STEIN \and D. M. YOUNG}
\affil{Department of Statistical Science, Baylor University, Waco, Texas 76706, U.S.A. \email{Caleb\_Stein@baylor.edu} \email{Dean\_Young@baylor.edu}}

@@

The popular Guo classifier is most competitive over large values of $d$ for the
Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
the implementation of the Guo method in the {\tt rda} package often classified
every test observation into a single class, thus yielding larger error rates
than its competitors. Furthermore, the Guo classifier exhibited unreliable
classification performance for smaller values of $d$ and required a larger
subset of the features to perform competitively in our study.
\begin{figure}
<<error_rates>>=
error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
p <- p + geom_line(aes(), size = 1, linetype = 2)
p <- p + geom_point(aes(shape = Classifier), size = 6)
p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
p <- p + theme(axis.text = element_text(size = 16))
p <- p + theme(axis.title.x = element_text(size = 20))
p <- p + theme(axis.title.y = element_text(size = 20))
p <- p + theme(strip.text.x = element_text(size = 18))
p + theme(legend.position = "none")
@ 
\caption{Averages of the 1000 error-rate estimates of the competing classifiers
  for each data set as a function of $d$. The classifiers are marked as follows:
  Dudoit (circle), Guo (triangle), Pang (square), Tong = (cross), Witten (
  square with X), and ours (asterisk). Approximate standard errors are no
  greater than $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.}
\label{fig:error-rates}
\end{figure}
Next, we examine the optimal estimates of the tuning parameters for our
classifier. For brevity, we provide graphical summaries of the optimal estimates
of $\lambda$ and $\gamma$ in the Supplementary Material. For each of the four
data sets, the distribution of the optimal values of $\lambda$ selected through
cross-validation often had a notable bathtub shape, resembling a Beta
distribution with both shape parameters less than unity. Specifically, the
optimal values of $\lambda$ were typically near 0 or near 1, corresponding to
quadratic and linear discriminant analysis, respectively. For instance, with
smaller values of $d$ applied to the Shipp data set, the range of optimal values
of $\lambda$ was typically no larger than 0.05, indicating there were a large
number of training data sets for which we can safely assume the covariance
matrices are unequal or nearly unequal. As $d$ increased, the majority of the
optimal values of $\lambda$ were selected as unity, thereby an increased number
of variables resulted in our classifier preferring the more parsimonious
assumption of equal covariance matrices. In this instance we can see the
flexibility of our classifier to employ the more realistic assumption that the
class covariance matrices are unequal, but if optimal, we can employ the more
parsimonious assumption of equal covariance matrices.

@@

\section*{Supplementary material}
\label{SM}
Supplementary material available at \Bka\ online includes the graphical summary
of the optimal estimates of our classifier from the classification study.

\bibliographystyle{biometrika}
\bibliography{rda}

>>>>

\documentclass[11pt]{article}
\usepackage{graphicx, amsmath, amssymb, bm, url, mathtools, natbib, amsthm, setspace}

\pagestyle{plain}
%----------------Page dimensions ----------------
\oddsidemargin 0.0in
\evensidemargin 0.0in
\topmargin -0.75in
\leftmargin 0in
\headheight 0.0in
\headsep 0.5in
%\footheight 0.0in
\footskip 0.5in
\footnotesep 0.0in
\textwidth 6.7in
\textheight 9.5in
%-----------Define Pictures---------------------
\def\picture #1 by #2 (#3){
   \vbox to #2{
        \hrule width #1 height 0pt depth 0pt
           \vfill
              \special{picture #3} % this is the low-level interface
                 }
    }
\def\scaledpicture #1 by #2 (#3 scaled #4){{
     \dimen0=#1 \dimen1=#2
      \divide\dimen0 by 1000 \multiply\dimen0 by #4
       \divide\dimen1 by 1000 \multiply\dimen1 by #4
        \picture \dimen0 by \dimen1 (#3 scaled #4)}
   }

\title{High-Dimensional Regularized Discriminant Analysis}

\author{John A. Ramey, Caleb K. Stein, and Dean M. Young}
%Vaccine and Infectious Disease Division, Fred Hutchinson Cancer Research
%  Center,\\ Seattle, Washington 98115, U.S.A. \email{johnramey@gmail.com}}
%\affil{Department of Statistical Science, Baylor University, Waco, Texas 76706, U.S.A. \email{Caleb\_Stein@baylor.edu} \email{Dean\_Young@baylor.edu}}

\newcommand{\xbar}{\bar{\bm x}}
\newcommand{\tr}{\text{tr}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\newtheorem{cor}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{thm}{Theorem}

@@

@@

\doublespacing

@@

The popular Guo classifier is most competitive over large values of $d$ for the
Shipp and Singh data sets. However, for the Burczynski and Nakayama data sets,
the implementation of the Guo method in the {\tt rda} package often classified
every test observation into a single class, thus yielding larger error rates
than its competitors. Furthermore, the Guo classifier exhibited unreliable
classification performance for smaller values of $d$ and required a larger
subset of the features to perform competitively in our study.
\begin{figure}
<<error_rates>>=
error_rates$data_set <- Hmisc:::capitalize(error_rates$data_set)
p <- ggplot(error_rates, aes(x = d, y = Avg_Error, group = Classifier))
p <- p + geom_line(aes(), size = 1, linetype = 2)
p <- p + geom_point(aes(shape = Classifier), size = 6)
p <- p + facet_wrap(~ data_set) + ylim(c(0, max(error_rates$Avg_Error))) + theme_bw()
p <- p + xlab("Number of Variables Selected, d") + ylab("Average Error Rate")
p <- p + theme(axis.text = element_text(size = 16))
p <- p + theme(axis.title.x = element_text(size = 20))
p <- p + theme(axis.title.y = element_text(size = 20))
p <- p + theme(strip.text.x = element_text(size = 18))
p + theme(legend.position = "none")
@ 
\caption{Averages of the 1000 error-rate estimates of the competing classifiers
  for each data set as a function of $d$. The classifiers are marked as follows:
  Dudoit (circle), Guo (triangle), Pang (square), Tong = (cross), Witten (square
  with X), and ours (asterisk). Approximate standard errors are no greater than
  $\Sexpr{gsub('.', '\\cdot', round(max_se, 3), fixed = TRUE)}$.}
\label{fig:error-rates}
\end{figure}
Next, we examine the optimal estimates of the tuning parameters for our
classifier. For brevity, we provide graphical summaries of the optimal estimates
of $\lambda$ and $\gamma$ in the Supplementary Material. For each of the four
data sets, the distribution of the optimal values of $\lambda$ selected through
cross-validation often had a notable bathtub shape, resembling a Beta
distribution with both shape parameters less than unity. Specifically, the
optimal values of $\lambda$ were typically near 0 or near 1, corresponding to
quadratic and linear discriminant analysis, respectively. For instance, with
smaller values of $d$ applied to the Shipp data set, the range of optimal values
of $\lambda$ was typically no larger than 0.05, indicating there were a large
number of training data sets for which we can safely assume the covariance
matrices are unequal or nearly unequal. As $d$ increased, the majority of the
optimal values of $\lambda$ were selected as unity, thereby an increased number
of variables resulted in our classifier preferring the more parsimonious
assumption of equal covariance matrices. In this instance we can see the
flexibility of our classifier to employ the more realistic assumption that the
class covariance matrices are unequal, but if optimal, we can employ the more
parsimonious assumption of equal covariance matrices.

@@

\section*{Supplementary material}
\label{SM}
Supplementary material available online includes the graphical summary
of the optimal estimates of our classifier from the classification study.

\bibliographystyle{plainnat}
\bibliography{rda}
